---
params:
  section: "abc"
  pg: "abc"
  var: "abc"
title: "CCG level summary scores"
output: 
 html_document: 
   self_contained: no
   mathjax: null
   highlight: null
   fig_width: 10
   theme: null
   template: "../CCO_template_html.html"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F,warning=F,message=F)

setwd("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/summary scores")

```

<link rel="stylesheet" href="W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis//styles.css" type="text/css" />

<a href="../../index.html">Home</a></li>

```{r import packages}

require(data.table)
require(dplyr)
require(tidyr)
require(knitr)
require(kableExtra)
require(downloadthis)

```

<script src="extra_js.js"></script>

<br/>
As with last year’s report, to provide an overall indication of how each CCG compares to the rest of England in terms of children’s access to mental health services, we created a summary score for each CCG based on five key indicators of CYPMHS performance. These indicators are:

1. CCG spending on children’s mental health as a percentage of total CCG allocation 
2. Mental health spend per child - calculated using NHS FYFVMH spending figures and ONS mid 2020 CCG population estimates
3. Total number of children referred to children’s mental health services as a proportion of the under-18 population.
4. Average waiting time for children who receive a second contact.
5. The percentage of referrals that are closed before treatment.

<br/>
For each indicator, CCGs are ranked from best to worst (e.g. shortest waiting time to longest) and assigned to 5 quintiles. Scores ranging from 1 to 5 are given to the CCG based on their quintile group. Best performing CCGs are given a score of 5 while the worst performers are given a score of 1. We then add these quintile scores together to form an overall score ranging from 5 to 25 for each CCG. An overall score of 5 would indicate being in the bottom quintile across all 5 measures while a score of 25 would indicate being in the top quintile across all measures.

***

**Overall CCG scores**

According to these scoring criteria, NHS Wakefield, NHS Tees Valley and NHS Heywood, Middleton and Rochdale were the best performers this year with overall scores of 25, 24 and 23 respectively.  

Conversely, NHS West Sussex and NHS Coventry and Rugby received the lowest overall scores of eight out of 25.



**Table 1. Top 20 best performing CCGs in 2020/21**

```{r 20 best, echo=FALSE}

###top CCGs ###

top_ccgs <- fread("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/tables/t1_ccg_top20.csv")

kable(top_ccgs) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

top_ccgs %>%
  download_this(
    output_name = "big ask data download",
    output_extension = ".xlsx",
    button_label = "Download data as xlsx",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save"
  )

```

<br/>

**Table 2. Bottom 20 CCGs with the lowest overall scores in 2020/21.**

```{r 20 worst, echo=FALSE}

###top CCGs ###

bottom_ccgs <- fread("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/tables/t2_ccg_bottom20.csv")

kable(bottom_ccgs) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

bottom_ccgs %>%
  download_this(
    output_name = "big ask data download",
    output_extension = ".xlsx",
    button_label = "Download data as xlsx",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save"
  )


```
