---
params:
  section: "abc"
  pg: "abc"
  var: "abc"
title: "Trends in children’s access to CYPMHS"
output: 
 html_document: 
   self_contained: no
   mathjax: null
   highlight: null
   fig_width: 10
   theme: null
   template: "../CCO_template_html.html"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F,warning=F,message=F)

setwd("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/children referred")

```

<link rel="stylesheet" href="W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis//styles.css" type="text/css" />

<a href="../../index.html">Home</a></li>


```{r import packages}

require(data.table)
require(dplyr)
require(tidyr)
require(knitr)
require(kableExtra)
require(downloadthis)

```

<script src="extra_js.js"></script>

### Number of children referred to CYPMHS

1. The number and rate of children referred to CYPMHS has decreased for the first time since 2017/18. Access rates (as calculated by the NHS) have increased, but these figures are based on 2004 mental health disorder prevalence (1 in 10 children), not the much higher 2021 prevalence estimates (1 in 6 children). 

*Note: All findings below need to be considered within the context of the Coronavirus pandemic. The crisis placed unprecedented pressure on the NHS’ capacity to deliver much-needed services. This, combined with the negative impact of social isolation on mental health, means that there are simultaneously more people in need of treatment while there are fewer resources to do so.*

* The number and rate of children referred to CYPMHS has decreased for the first time since 2017/18, the year when the OCC published the first iteration of this mental health briefing. In 2020/21, 497,502 children in England were referred to CYPMHS. This is equal to 4% of all children in England – a decrease from 4.5% in 2019/20 but an increase from 3.4% in 2018/19.

* This year, only 31 CCGs (23%) saw positive changes in the percentage of children referred to CYPMHS. Upon closer inspection, none of these CCGs saw large increases in the rate of children referred. The largest percentage point increase (shared by four CCGs) was 1.1%. The CCGs are as follows: NHS Isle of Wight, NHS Northumberland, NHS Nottingham and Nottinghamshire and NHS Wirral.

* The remaining 104 CCGs (77%), saw decreases in the number referred to CYPMHS – the largest of which were seen in NHS Barking and Dagenham (-2.2 ppts) and NHS Stafford and Surrounds (-2.0 ppts).


**Table 1. The 10 CCGs with the largest increases in percentage of children referred to CYPMHS**

```{r referrals per year, echo=FALSE}

# READ IN TABLE

### Largest increases in perc referred ###

percreferred <- fread("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/tables/t1_percreferred.csv")

kable(percreferred) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

percreferred %>%
  download_this(
    output_name = "big ask data download",
    output_extension = ".xlsx",
    button_label = "Download data as xlsx",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save"
  )

```
<br/>

Access rates (as calculated by the NHS) have continued to increase from an average of 36.8% in 2019/20 to 39.6% in 2020/21. However, these figures are based on 2004 mental health disorder prevalence which does not take into account the impact of the pandemic on children’s mental health. New data published in the 2021 update to the NHS Mental Health of Children and Young People report show that the rates of children with probable mental disorders have increased since 2017: 

* For 6 to 16 year olds from one in nine (11.6%) to one in six (17.4%). 
* For 17 to 19 year olds from one in ten (10.1%) to one in six (17.4%).

Applying these new prevalence rates to ONS mid-2020 population estimates for those aged 6 to 17 shows that there is a pool of 1,313,983 children with a probable mental health disorder. Data from the NHS FYFV dashboard reports that there were 420,504 young people in 2020/21 who received support from NHS funded community services. From this we calculate that the true access rate is closer to 32%, a notable drop from 39.6% calculated based on 2004 prevalence figures.


**Table 2. Change in national access rates from 2017/18 to 2019/20**

```{r accesstrend, echo=FALSE}

# READ IN TABLE

### Changes in access rates ###

accesschange <- fread("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/tables/t2_accesschange.csv")

accesschange <- accesschange %>% 
  select(!V5)

kable(accesschange) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

accesschange %>%
  download_this(
    output_name = "big ask data download",
    output_extension = ".xlsx",
    button_label = "Download data as xlsx",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save"
  )

```
<br/>

*Note: Access rate figures from the FYFV dashboard were marked as experimental until 2019/20. As a result, exercise caution in interpreting changes over time. Figures for the number of children accessing treatment were not released until 2020/21, therefore access rates (2020/21 prevalence) for the years before are not calculable.*


