---
title: "Briefing on Children’s Mental Health Services – 2020/2021"
output: 
 html_document: 
   self_contained: no
   theme: null
   mathjax: null
   highlight: null
   css: styles.css
   template: "CCO_template_html.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)


setwd("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/")
```

<link rel="stylesheet" href="../../styles.css" type="text/css" />


```{r}

#<link rel="stylesheet" href="../../styles.css" type="text/css" />

```
**Introduction**

This is the Office of the Children’s Commissioner’s (OCC) fifth annual report on the state of children’s
mental health services in England. The report aims to assess children’s ability to access Children and
Young People’s Mental Health Services (CYPMHS) provision in England in 2020/21 and how (if at all) this
has changed over the past two years. Using new figures sourced from NHS Digital, NHS England and
NHS Improvement, this report examines the following at both a national and Clinical Commissioning
Group (CCG) level:

* Numbers of children referred to and accessing CYPMHS
* Numbers of children referred that were not accepted into treatment during the 2020/21 financial year, as proxied by the numbers not receiving two contacts with CYPMHS
* Average waiting times
* Spending on children’s mental health

</br>

**Executive Summary**

* **Nationally, spending on specialist NHS mental health services for children has increased for the fourth consecutive year across the majority of CCGs.** In 2020/21, £881 million was spent on children’s mental health services (an increase from £791 million the previous year) which is an increase of 11% in cash, and 4.4% in real terms. Out of 135 CCGs in England – 111 (82%) increased their spending on mental health from last year. 

* **Half of CCGs spend at least 1% of their budget on CYPMHS. This is the NHS Long Term Plan benchmark of how much CCGs should be spending on CYPMHS.** 68 out of 135 CCGs (50%) spent at least 1% of their budget on CYPMHS. This is an improvement from 2018/19 where 59 out of 195 CCGs (30%) spent at least 1% of their budget on CYPMHS, and from 2019/20 where 69 out of 191 CCGs (36%) reached that 1% target.

* **The number of children referred (for example by their GP or teacher) to NHS children’s mental health services has decreased for the first time in four years. In 2020/21, 497,502 children were referred, a decrease from 539,000 the previous year.** This could be due to disruption caused by the pandemic. The percentage of children being referred nationally has also decreased. Referral rates have dropped from 4.5% to 4% of the under-18 population.

* **More children have been struggling with their mental health since 2017 – which covers the pandemic period.** One in six children have a probable mental health disorder. This is up from one in nine children with a probable mental health disorder in 2017. While progress has been made to reduce the treatment gap between those who need treatment and what is provided, this increase in need has made this more challenging.

* **Overall, progress has been made across all of our indicators: on waiting times and a reduction in the number of children turned away.**

* **For those children who are accepted into mental health treatment services, average waiting times have decreased substantially from last year.** The waiting times have decreased from 43 days in 2019/20 to 32 days in 2020/21. 

* **However, over a third (37%) of children accepted onto waiting lists are still waiting for their treatment to begin.**


</br>

**Data sources and methodology**

<div class="topic_navList">
<a class="meta_navList_link" href="data sources/pages/data_sources.html">
<h3 class="list_header">Data sources</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="limitations/pages/limitations.html">
<h3 class="list_header">Limitations of data and analysis</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="Methodology/pages/methodology.html">
<h3 class="list_header">Methodology</h3></a>
</div>

**Trends in children's access to CYPMHS**

<div class="topic_navList">
<a class="meta_navList_link" href="children referred/pages/children-referred.html">
<h3 class="list_header">Number of children referred to CYPMHS</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="referrals closed/pages/referrals-closed.html">
<h3 class="list_header">Referrals closed before treatment</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="waiting times/pages/waiting-times.html">
<h3 class="list_header">Waiting times and outcomes</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="spend/pages/spend.html">
<h3 class="list_header">CCG spending on children's mental health services</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="summary scores/pages/summary-scores.html">
<h3 class="list_header">CCG summary scores</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="ccg maps/pages/ccg-maps.html">
<h3 class="list_header">Local area (CCG) maps</h3></a>
</div>


