---
title: "The state of children's mental health services 2020/21"
output: 
 html_document: 
   self_contained: no
   theme: null
   mathjax: null
   highlight: null
   css: styles.css
   template: "CCO_template_html.html"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)


setwd("W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis/")
```

<link rel="stylesheet" href="W:/CCO-WORKING-FS/Projects/MH briefing/MH briefing 202021/analysis//styles.css" type="text/css" />


```{r}

#<link rel="stylesheet" href="../../styles.css" type="text/css" />

```
**Introduction**

This is the Office of the Children’s Commissioner’s (OCC) fifth annual report on the state of children’s
mental health services in England. The report aims to assess children’s ability to access Children and
Young People’s Mental Health Services (CYPMHS) provision in England in 2020/21 and how (if at all) this
has changed over the past two years. Using new figures sourced from NHS Digital, NHS England and
NHS Improvement, this report examines the following at both a national and Clinical Commissioning
Group (CCG) level:

* Numbers of children referred to and accessing CYPMHS
* Numbers of children referred that were not accepted into treatment during the 2020/21 financial year, as proxied by the numbers not receiving two contacts with CYPMHS
* Average waiting times
* Spending on children’s mental health

</br>

**Headline findings**

* Fewer children are being referred to CAMHS – in 2020/21, 484,000 children were referred, a decrease from 539,000 last year. Referral rates have also dropped from 4.5% to 4% of the under 18 population.
* For those who are accepted into treatment, average waiting times have decreased substantially from 43 days (2019/20) to 32 days (2020/21). 
* However, many children are not getting the help they need. Over a third of children referred are still waiting for their second contact (up from last year) and almost a quarter had their referrals closed before treatment.
* Spending has increased for the fourth consecutive year. In 2020/21, £881 million was spent on children’s mental health services – an increase from £791 million last year.

</br>

**Data sources and methodology**

<div class="topic_navList">
<a class="meta_navList_link" href="data sources/pages/data_sources.html">
<h3 class="list_header">Data sources</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="limitations/pages/limitations.html">
<h3 class="list_header">Limitations of data and analysis</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="Methodology/pages/methodology.html">
<h3 class="list_header">Methodology</h3></a>
</div>

### **Trends in children's access to CYPMHS**

<div class="topic_navList">
<a class="meta_navList_link" href="children referred/pages/children referred.html">
<h3 class="list_header">Number of children referred to CYPMHS</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="referrals closed/pages/referrals closed.html">
<h3 class="list_header">Referrals closed before treatment</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="waiting times/pages/waiting times.html">
<h3 class="list_header">Waiting times and outcomes</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="spend/pages/spend.html">
<h3 class="list_header">CCG spending on children's mental health services</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="summary scores/pages/summary scores.html">
<h3 class="list_header">CCG summary scores</h3></a>
</div>

<div class="topic_navList">
<a class="meta_navList_link" href="ccg_maps/pages/ccg_maps.html">
<h3 class="list_header">Local area (CCG) maps</h3></a>
</div>


